# Chapter 6
# RATS datas

```{r, include=FALSE}
RATSL <- read.table("RATSL.txt")
BPRSL <- read.table("BPRSL.txt")
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep = "\t", header = T)
RATSL$ID <- as.factor(RATSL$ID)
RATSL$Group <- as.factor(RATSL$Group)
BPRSL$treatment <- as.factor(BPRSL$treatment)
BPRSL$subject <- as.factor(BPRSL$subject)
```

**Importing libraries**
```{r}
library(dplyr)
library(tidyr)
library(lme4)
library(ggplot2)
```

**Looking at data of RATS**
```{r}
glimpse(RATSL)
```
As we can see, the RATS data has 176 rows and 5 columns. If we look at the data sources, we can see that there are
16 rats and three groups. Eight rats are in group one, four rats are in group 2, and the last four are in group 3.


**Visualizating the data**
```{r}
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))
```
If we look at the plot, we can see that the rats from group 1 weight the least, but the other groups have increased weight, but we need to remember that groups two and three weighted more in the beginning of the tests.

**Standardizing weights and visualizating new plot**
```{r}
RATSL <- RATSL %>%
  group_by(Group) %>%
  mutate(stdweight=scale(Weight) ) %>%
  ungroup()

glimpse(RATSL)
```
```{r}
# Plot the data
ggplot(RATSL, aes(x = Time, y = stdweight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "Standardized Weight") +
  theme(legend.position = "none")
```

# BPRS data
The BPRS data contains a study that has 40 subjects in two different groups with "Brief Psychiatric Rating Scale" aka BPRS.

**Looking at the data:**
```{r}
glimpse(BPRSL)
```

**Visualizating the data**
```{r}
ggplot(BPRSL, aes(x = week, y = bprs, group = subject)) +
  geom_line(aes(linetype=subject)) + facet_grid(. ~ treatment, labeller = label_both) +
  scale_x_continuous(name= "Weeks") + scale_y_continuous(name = "BPRS") + theme(legend.position = "top") +
  theme(legend.position = "none")
```

**Fitting the data to two models**
```{r}
model1 <- lmer(bprs ~ week + treatment + (weeks | subject), data = BPRSL)

model2 <- lmer(bprs ~ week * treatment + (weeks | subject), data = BPRSL)

summary(model1)
summary(model2)
```

**Comparing the two fitted models**
```{r}
anova(model1, model2)
```
By looking at the models, according to anova models we can see that there is practically no significant differences between the two treatment groups.